A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-05-23 discussion.tex
top 10

Furthermore, we expected to obtain some patterns associated with certain Rust-related features. However, our results did not include those patterns. For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to trouble many beginning Rust programmers. \verb+move+ is used to transfer the ownership of values to a closure. In some cases (e.g. multithreading), not moving the ownership could create invalid references. Rust compiler checks prevent such errors. Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many errors, and they only push their code once it is free of compile errors. We thus would not observe patterns which don't make it to the repository.

@@

We believe our results can help researchers to create IDE tools for Rust. We have discussed the actionability of some of the general patterns and three BC-related patterns that are not detected by Clippy. For instance, we discussed considering the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could be created to search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere. Also, the linter can analyze the usages of the struct to determine the necessity of adding the attribute. Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces the binary size. Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}. As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive and might lead to hogging the CPU. Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing. These findings and insights can greatly aid researchers in creating effective IDE tools tailored for Rust development. Moreover, the empirical frequency numbers associated with each pattern, highlighting its prevalence, can provide valuable guidance for the development of program repair tools.

>>>>

\subsection{Patterns Present and Missing}

When presenting each pattern, we discussed actionability for the most important actions. Here, we discuss the implications of our most important patterns.

The two most common patterns were general patterns modifying struct fields and attributes, with 179 and 172 instances respectively. Unfortunately, struct field modifications are neither interesting nor actionable; they are simply used to enable other changes. Attribute modifications are far more interesting, and account for the second- and fifth-most popular pattern. From a language design perspective, the fact that there are so many attribute changes marked as bug fixes illustrates the semantic importance of attributes in Rust, and implies that program repair must understand attributes. A further challenge is that different attributes have different, and often library-dependent, semantic meanings. Proposing specific repair techniques that reason about attributes is beyond the scope of this work, but our results suggest that they are important to develop.

The most common borrow-checker-related pattern and third-most-common pattern overall is dropping \texttt{clone()}. This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption. Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues. This implies that program repair tools would benefit from being performance-aware, rather than limiting themselves strictly to functional properties of the code.

The fourth- and sixth-most popular patterns, changing the type of a struct field or of a generic type parameter, are difficult to make general statements about; there are many unrelated idiosyncratic changes in these clusters. Such changes look the same, but have different purposes. One example we showed changed a \texttt{HashMap} to an \texttt{IndexMap}, preserving iteration order. That specific nondeterminism bug could be automatically repaired upon direction from the developer. Another change modified \texttt{usize} to \texttt{isize}, perhaps reflecting a requirements change to allow signed integers. IDEs could support refactorings that modify types. We find it difficult to propose further general work in this direction.

The seventh-most-popular pattern is the Rust-specific change of dropping mutability (i.e. making data immutable). Tools like ReImInfer~\cite{HuangMDE2012} for a Java extension, and techniques like the one proposed by Eyolfson~\cite{EyolfsonAbstractImmutability} for C++, aim to infer that data is immutable. In the Rust context, it is easier to make data immutable: the semantics for immutability are clear. Immutability is part of the core Rust language. mmutability escape hatches in C++ complicate issues, but in Rust, the escape hatch is the clearly-marked \texttt{unsafe} block. This pattern thus suggests that tools for inferring Rust immutability could be viable.

We are unable to say anything meaningful about the eighth- and ninth-most popular patterns, \texttt{G.match.pattern} and \texttt{G.type.enum.variant}. They do not seem particularly helpful for automatic program repair.

The tenth-most-popular pattern, making a struct field \texttt{pub} and hence visible to external modules, suggests that tools to help developers manage visibility would be helpful. Public accessibility of methods and fields very much affects library compatibility and usability.

Conversely, we were surprised to find that Ruxanne reported no patterns associated with certain Rust-related features. For instance, we did not observe any patterns associated with the Rust \verb+move+ operator, which has been known to trouble many beginning Rust programmers. \verb+move+ is used to transfer the ownership of values to a closure. In some cases (e.g. multithreading), not moving the ownership could create invalid references. Rust compiler checks prevent such errors. Our speculation is that, since Rust enforces strict compiler checks, programmers usually get many compile-time errors, but they only push their code once it is free of compile errors. We thus would not observe patterns which do not make it to the repository.

@@

We believe our results can help with the development of IDE and program repair tools for Rust. We have discussed the actionability of some of the general patterns and three BC-related patterns that are not detected by Clippy. For instance, as we have discussed, for the common pattern of \textit{Modifying the attributes of structs}, a code linter tool could search for correct attribute lists within the codebase~\citep{forrest2009genetic}, leveraging the assumption that proper attribute usage can be found elsewhere. Also, the linter can analyze the usages of the struct to determine the necessity of adding the attribute. Removing unnecessary attributes not only optimizes the codebase but also eliminates unnecessary overhead and reduces binary size. Similarly, we presented the pattern \textit{Dropping clone and adding borrowing}. As discussed, Clippy does not detect this pattern, and repetitive clones can be computationally expensive. Using state-of-the-art methods for program repair, one could design a tool to recognize this pattern and change variable cloning to simple borrowing. These findings and insights can greatly aid researchers in creating effective IDE tools tailored for Rust development. Moreover, the empirical frequency numbers associated with each pattern, highlighting its prevalence, can provide valuable guidance for the development of program repair tools.

### 2023-05-23 results.tex
top 10

@@

\noindent\textbf{Actionability:} The patterns related to access modifiers offer an opportunity for repair tools to make sure that the reachability of program entities follows the user's specifications, enhancing security and maintainability. Additionally, the empirical numbers revealing the frequency of struct field modifications in Rust shed light on the significance of these bug fix patterns and emphasize the need for further exploration on how program repair tools can effectively modify code based on changes in struct fields to repair program behavior. These patterns provide valuable insights for program repair tools to automatically analyze and modify the codebase, ensuring it aligns with the desired behavior.

@@

\noindent\textbf{Description:} In this pattern, a change occurs in an arm's pattern. Like the code snippet provided above, it can be a change in pattern's type. It also can be matching to a different enum variant or string. The fix pattern can happen both for refactoring or debugging purposes. \\

@@

\noindent\textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints. Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes. There are more than 500 lints included in Clippy. Since detecting whether a clone is redundant is undecidable, Clippy would have to conservatively estimate redundancy. By definition, removing a redundant clone does not modify functional aspects of the program behaviour. However, such a removal can greatly improve performance, if the program was formerly cloning large objects. \\

@@

\noindent\textbf{Actionability:} A program repair tool can efficiently implement this modification by examining the necessity of the clone operation and considering the user's specifications regarding variable cloning. If the clone is deemed unnecessary, the tool can effortlessly replace it with a reference to reduce the overhead of unnecessary copying, as discussed in the example above. \\

>>>>

\todo{please confirm: G.attr.struct.field is not a subset of G.attr.struct.}

@@

\noindent\textbf{Actionability:} The patterns related to access modifiers suggest an opportunity for repair tools to make sure that the accessability of program entities follows the user's specifications, enhancing encapsulation, security, and maintainability. Additionally, the empirical numbers revealing the frequency of struct field modifications in Rust shed light on the significance of these bug fix patterns and emphasize the need for further exploration on how program repair tools can effectively modify code based on changes in struct fields to repair program behavior.

@@

\noindent\textbf{Description:} In this pattern, a change occurs in an arm's pattern. Like the code snippet provided above, it can be a change in a match pattern's type. It also can be matching to a different enum variant or string. The fix pattern can happen both for refactoring or debugging purposes. \\

@@

\noindent\textbf{Description:} This pattern is referred to as redundant clone in Clippy Lints. Clippy\footnote{\url{https://github.com/rust-lang/rust-clippy}} is a linter tool for Rust that catches common mistakes. There are more than 500 lints included in Clippy. Since detecting whether a clone is redundant is undecidable, Clippy conservatively estimates redundancy. By definition, removing a clone that is indeed redundant does not modify functional aspects of the program behaviour. (Removing a non-redundant clone introduces a compile error or, worse yet, a bug.) However, such a removal can greatly improve performance, if the program was formerly cloning large objects. \\

@@

\noindent\textbf{Actionability:} A program repair tool can implement clone removal by examining the necessity of the clone operation and considering the user's specifications regarding variable cloning. If the clone is unnecessary, the tool can replace it with a reference to reduce the overhead of unnecessary copying, as discussed in the example above. \\
